Towards human SuperEEG

### 2018-10-04 paper/main.tex

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity with perfect precision and at arbitrarily high
  spatiotemporal resolution.  Although true SuperEEG is impossible
  using existing methods, here we present a model-based method for
  \textit{inferring} neural activity at millimeter-scale spatial
  resolutions and millisecond-scale temporal resolutions using
  standard human intracranial recordings.  Our approach assumes that
  different people's brains exhibit similar spatial correlations, and
  that (all else being equal) neural activity at nearby locations will
  tend to be similar.  One can then ask, for an arbitrary individual's
  brain: given recordings from a limited set of locations in that
  individual's brain, along with the observed spatial correlations in
  other people's recordings, what would recordings most likely have
  looked like at \textit{other} locations in that individual's brain?
%\\
%\small{\textbf{Keywords: Electrocorticography (ECoG), intracranial electroencephalography (iEEG),
%local field potential (LFP), epilepsy, maximum likelihood estimation, Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g.\ as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g.\ as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high temporal and spatial resolutions of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

%Beamforming: huge advance for scalp recordings, but the inverse
%problem is fundamentally ambiguous.  This limits the precision with
%which one can measure activity from deep brain structures.
An increasingly popular approach is to improve the effective spatial
resolution of MEG or scalp EEG data by using a geometric approach
called \textit{beamforming} to solve the biomagnetic or bioelectrical
inverse problem~\cite{Sarv87}.  This approach entails using detailed
brain conductance models (often informed by high spatial resolution
anatomical MRI images) along with the known sensor placements
(localized precisely in 3D space) to reconstruct brain signals
originating from theoretical point sources deep in the brain (and far
from the sensors).  Traditional beamforming approaches must overcome
two obstacles.  First, the inverse problem beamforming seeks to solve
has infinitely many solutions.  Researchers have made traction towards
constraining the solution space by assuming that signal-generating
sources are localized on a regularly spaced grid spanning the brain
and that individual sources are small relative to their distances to
the sensors~\cite{Snyd91, BailEtal01, HillEtal05}.  The second, and in
some ways much more serious, obstacle is that the magnetic fields
produced by external (noise) sources are substantially stronger than
those produced by the neuronal changes being sought (i.e.\ at deep
structures, as measured by sensors at the scalp).  This means that
obtaining adequate signal quality often requires averaging the
measured responses over tens to hundreds of responses or trials
(e.g. see review by~\cite{HillEtal05}).

Another approach to obtaining high spatial and temporal resolution
neural data has been to collect fMRI and EEG data simultaneously.
Simultaneous fMRI-EEG has the potential to balance the high spatial
resolution of fMRI with the high temporal resolution of scalp EEG,
thereby, in theory, providing the best of both worlds.  In practice,
however, the signal quality of both recordings suffers substantially
when the two techniques are applied simultaneously (e.g. see review
by~\cite{HustEtal12}).  In addition, the experimental designs that are
ideally suited to each technique individually are somewhat at odds.
For example, fMRI experiments typically lock stimulus presentation
events to the regularly spaced image acquisition time (TR), which
maximizes the number of post-stimulus samples.  By contrast, EEG
experiments typically employ jittered stimulus presentation times to
maximize the experimentalist's ability to distinguish electrical brain
activity from external noise sources such as from 60 Hz alternating
current power sources.

%Direct intracranial recordings are the gold standard method for
%measuring activity from deep brain structures in both humans and
%animals.
%However, one can only implant a limited number of electrodes in a
%given person's brain.  This begs the question: what can we infer
%about the activity exhibited by the rest of that person's brain,
%given what we learn from recordings taken from \textit{other people's
%brains}?
The current ``gold standard'' for precisely localizing signals and
sampling at high temporal resolution is to take (ECoG or iEEG)
recordings from implanted electrodes (but from a limited set of
locations in any given brain).  This begs the following question: what
can we infer about the activity exhibited by the rest of a person's
brain, given what we learn from the limited intracranial recordings we
have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we
call \textit{SuperEEG}, based on Gaussian process
regression~\cite{Rasm06}.  SuperEEG entails using data from multiple
people to estimate activity patterns at arbitrary locations in each
person's brain (i.e., independent of their electrode placements).  We
test SuperEEG approach using two large datasets of intracranial
recordings~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12, EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  We show
that the SuperEEG algorithm recovers signals well from electrodes that
were held out of the training dataset.  We also examine the factors
that influence how accurately activity may be estimated (recovered),
which may have important implications for electrode design and
placement in neurosurgical applications.

@@

@@

The first dataset was taken from 6876 electrodes implanted in the
brains of 88 epilepsy patients~\cite{SedeEtal03, SedeEtal07a,
  SedeEtal07b, MannEtal11, MannEtal12}.  We first applied a fourth
order Butterworth notch filter to remove 60 Hz ($\pm$ .5 Hz) line
noise.  We then excluded any electrodes that showed putative
epileptiform activity.  Specifically, we excluded from further
analysis any electrode that exhibited an average kurtosis of 10 or
greater across all of that patient's recording sessions.  We also
excluded any patients with fewer than 2 electrodes that passed this
criteria, as the SuperEEG algorithm requires measuring correlations
between 2 or more electrodes from each patient.  Altogether this
yielded clean recordings from 4168 electrodes implanted throughout the
brains of 67 patients (Fig.~\ref{fig:methods}A). For the purposes of
comparing task-specific contributions to reconstruction accuracy, we
limited our analyses in the second dataset to patients that
participated in two free recall experiments.  Applying the same
kurtosis thresholding yielded clean recordings from 24 patients and
2975 electrodes for the second dataset. Each individual patient
contributes electrodes from a limited set of brain locations, which we
localized in a common space (MNI152); an example patient's 54
electrodes that passed the predefined kurtosis test are highlighted in
black and red.

The recording from a given electrode is maximally informative about
the activity of the neural tissue immediately surrounding its
recording surface.  However, brain regions that are distant from the
recording surface of the electrode also contribute to the recording,
albeit (often) to a much lesser extent.  One mechanism underlying these
contributions is volume conduction.  The precise rate of falloff due
to volume conduction (i.e. how much a small volume of brain tissue at
location $x$ contributes to the recording from an electrode at
location $\eta$) depends on the size of the recording surface, the
electrode's impedance, and the conductance profile of the volume of
brain between $x$ and $\eta$.  As an approximation of this intuition,
we place a Gaussian radial basis function (RBF) at the location $\eta$
of each electrode's recording surface (Fig.~\ref{fig:methods}B).  We
use the values of the RBF at any brain location $x$ as a rough
estimate of how much structures around $x$ contributed to the
recording from location $\eta$:

\begin{align}
  \mathrm{rbf}(x|\eta,\lambda) & =
  \mathrm{exp}\left\{ -\frac{||x - \eta||^2}{\lambda} \right\},\label{eqn:rbf}
\end{align}
where the width variable $\lambda$ is a parameter of the algorithm
(which may in principle be set according to location-specific tissue
conductance profiles) that governs the level of spatial smoothing.  In
choosing $\lambda$ for the analyses presented here, we sought to
maximize spatial resolution (which implies a small value of $\lambda$)
while also maximizing the algorithm's ability to generalize to any
location throughout the brain, including those without dense electrode
coverage (which implies a large value of $\lambda$). Using our prior
work as a guide~\cite{MannEtal14b}, we set $\lambda = 20$, although
this could in theory be optimized, e.g. using cross validation.

A second mechanism whereby a given region $x$ can contribute to the
recording at $\eta$ is through anatomical connections between
structures near $x$ and $\eta$.  We use spatial correlations in the
data to estimate these anatomical connections.  Let $\bar{R}$ be the
set of locations at which we wish to estimate local field potentials,
and let $R_{s}$ be set of locations at which we observe local field
potentials from patient $s$ (excluding the electrodes that did not
pass the kurtosis test described above). In the analyses below we
define $\bar{R} = \cup_{s=1}^S R_s$.  We can calculate the expected
inter-electrode correlation matrix for patient $s$, where
$C_{s,k}(i,j)$ is the correlation between the time series of voltages
for electrodes $i$ and $j$ from subject $s$ during session $k$, using:

\begin{align}
  \bar{C}_{s} &=
  \mathrm{r}(\frac{1}{n}(\sum_{k=1}^{n}\mathrm{z}(C_{s,k}))),\label{eqn:inter_corr}~\mathrm{where}\\
\mathrm{z}(r) &= \frac{\log(1+r) - \log(1 - r)}{2}~\mathrm{is~the~Fisher}~z \mathrm{-transformation~and}\label{eqn:fishersz}\\
\mathrm{z}^{-1}(z) = \mathrm{r}(z) &= \frac{\exp(2z) - 1}{\exp(2z) + 1}\mathrm{~is~its~inverse}.\label{eqn:invfishersz}
\end{align}
Next, we use Equation~\ref{eqn:rbf} to construct a number of
to-be-estimated locations by number of patient electrode locations
weight matrix, $W$.  Specifically, $W$ approximates how informative the
recordings at each location in $R_s$ are in reconstructing activity at
each location in $\bar{R}$, where the contributions fall off with an RBF according to
the distances between the corresponding locations:

\begin{align}
W(i, j) = \mathrm{rbf}(i|j,\lambda)\label{eqn:weight_matrix}.
\end{align}

Given this weight matrix, $W$,
and the observed inter-electrode correlation matrix for patient $s$,
$\bar{C}_{s}$, we can estimate the correlation matrix for all locations in
$\bar{R}$ (Fig.~\ref{fig:methods}C) using:

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:correlation_matrix_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j). \label{eqn:correlation_matrix_den}
\end{align}
Intuitively, we construct an estimated correlation matrix from each
individual patient's data (Fig.~\ref{fig:methods}C) using
Equations~\ref{eqn:correlation_matrix_num} \& \ref{eqn:correlation_matrix_den}, then we sum across these estimates for $S$ patients and divide (see Equations~\ref{eqn:correlation_matrix}) to obtain the expected correlation
matrix, $\hat{K}$ (Fig.~\ref{fig:methods}D):

\begin{align}
 \hat{K} = \mathrm{r} \left (  \frac{\sum  \hat{N}_{s}}{\sum \hat{D}_{s}}\right ).\label{eqn:correlation_matrix}
\end{align}
Now we can use the following intuition: given (i) the observed
responses from a limited set of locations in $R_{s}$ ($Y_{s}$) and
(ii) how each location's responses relate to all other responses
($\hat{K}$), we can estimate the LFP data from patient
$s$, for any arbitrary location in $\bar{R}$
(Fig.~\ref{fig:methods}E).

Let $\alpha$ be the set of indices of patient $s$'s electrode locations in
$\bar{R}$, and let $\beta$ be the set of indices of all other
locations in $\bar{R}$. In other words, $\beta$ reflects the locations
in $\bar{R}$ where we did not observe a recording for patient $s$
(these are the recording locations we will want to fill in using SuperEEG).
We can sub-divide $\hat{K}$ as follows:

\begin{align}
\hat{K}_{\beta,\alpha} &= \hat{K}(\beta,\alpha),~\mathrm{and}\label{eqn:Kba}\\
\hat{K}_{\alpha,\alpha} &= \hat{K}(\alpha,\alpha)\label{eqn:Kaa}.
\end{align}
Here $\hat{K}_{\beta, \alpha}$ stores the correlations between the
``unknown'' activity at the locations in $\beta$ and the observed
activity at the locations in $\alpha$, and $\hat{K}_{\alpha, \alpha}$
stores the correlations between the observed recordings (at the
locations in $\alpha$).

Let $Y_{s,k,\alpha}$ be the number-of-timepoints
($T$) by length($\alpha$) matrix of (observed) voltages from the electrodes in $\alpha$ during session $k$ from patient $s$. Then we can estimate the voltage from patient $s$'s $k^{th}$ session at the locations in $\beta$ using~\cite{Rasm06}:

\begin{align}
Y_{s,k,\beta} &= ((\hat{K}_{\beta,\alpha}\cdot\hat{K}_{\alpha,\alpha}^{-1})\cdot Y_{s,k,\alpha}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas
we observe the recordings only at the locations in $\alpha$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations in $\beta$, which we can define \textit{a priori} to
include any locations we wish, throughout the brain.  This yields estimates of
the time-varying voltages at \textit{every} location in $\bar{R}$.

We designed our approach to be agnostic to electrode impedances, as
electrodes that do not exist do not have impedances.  Therefore our
algorithm recovers voltages in standard deviation ($z$-scored) units
rather than attempting to recover absolute voltages. (This property
reflects the fact that $\hat{K}_{\beta, \alpha}$ and
$\hat{K}_{\alpha, \alpha}$ are correlation matrices rather than
covariance matrices.)  Also, note that
Equation~\ref{eqn:reconstruction} requires computing a $T$ by $T$
matrix, which can become computationally intractable when $T$ is very
large (e.g.\ for the patient highlighted in Fig.~\ref{fig:recon}, $T = 20458799$). However, we may approximate $Y_{s,k,\beta}$ in a piecewise
manner by filling in $Y_{s,k,\beta}$ in blocks of size $b$ samples
(using the corresponding samples from $Y_{s, k, \alpha}$).  In our
computations we set $b = 25000$.

The SuperEEG algorithm described above and in
Figure~\ref{fig:methods} allows us to estimate (up to a constant
scaling factor) LFPs for each patient at all arbitrarily chosen
locations in the set $\bar{R}$, \textit{even if we did not record that
  patient's brain at all of those locations}.

@@

Holding out each electrode from each patient in turn, we computed the
average correlation (across recording sessions) between the Super
EEG-reconstructed voltage traces and the observed voltage traces from
that electrode.  For each reconstruction, we estimated the full-brain
correlation matrix using every \textit{other} patient's data
(i.e. every patient except the one who contributed the
to-be-reconstructed electrode data).  In our analyses, we then
substituted the average correlation matrix computed after excluding
patient $s$'s data for $\hat{K}$ in Equations~\ref{eqn:Kba} and
\ref{eqn:Kaa}.  This step ensured that the data we were reconstructing
could not also be used to estimate the between-location correlations
that drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).

@@

%why does it work?  redundancy of brain
A fundamental assumption of the SuperEEG algorithm is that the data
covariance matrix is stable over time and across people.  This is a
useful simplification.  However, a growing body of evidence from the
fMRI community suggests that the data covariance matrix changes in
meaningful ways over time (for example, the data covariance matrix
changes from moment-to-moment during story listening, serving as a
unique ``fingerprint'' for each moment of the story; further, these
task-driven timepoint-specific covariance fingerprints appear to be
largely preserved across people~\cite{SimoEtal16, MannEtal17}).  These
findings indicate that the full-brain covariance matrix is not stable
over time.  Other recent work has shown that people's resting state
connectivity matrices may be used to uniquely identify individuals and
predict fluid intelligence scores~\cite{FinnEtal15}.  This indicates
that the full-brain covariance matrix is not stable across people.  If
the fundamental stability assumptions that SuperEEG relies on are
violated, how can the SuperEEG algorithm still accurately recover LFP
data?  It is important to recognize that the fact that variability
(over time or across people) is predictive (e.g.\ of cognitive states
during story listening or fluid intelligence scores) does not
necessarily mean that this variability is large in magnitude.  Rather,
we have long known that brain structure is tightly preserved across
individuals (and over time, at least on the timescale of typical
clinical and experimental recording sessions), and any functional
changes must occur within the framework of the underlying structural
anatomy.  Nevertheless, one could imagine future improvements to the
SuperEEG approach that leverage resting state fMRI or structural data
[e.g. diffusion tensor imaging (DTI)] to estimate Bayesian priors over
the correlation matrices inferred, in the current framing, using only
ECoG data.  Further, relaxing the assumption that the covariance
matrix is stable (over time and/or across people), and/or
incorporating more detailed brain conductance models (e.g.\ informed
by structural MRI scans) may improve the predictive performance of the
approach.

>>>>

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity with perfect precision and at arbitrarily high
  spatiotemporal resolution.  Although true SuperEEG is impossible
  using existing methods, here we present a model-based method for
  \textit{inferring} neural activity at millimeter-scale spatial
  resolutions and millisecond-scale temporal resolutions using
  standard human intracranial recordings.  Our approach assumes that
  different people's brains exhibit similar spatial correlations, and
  that (all else being equal) neural activity at nearby locations will
  tend to be similar.  One can then ask, for an arbitrary individual's
  brain: given recordings from a limited set of locations in that
  individual's brain, along with the observed spatial correlations in
  other people's recordings, what would recordings most likely have
  looked like at \textit{other} locations in that individual's
  brain?\\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g., as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g., as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high temporal and spatial resolutions of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

An increasingly popular approach is to improve the effective spatial
resolution of MEG or scalp EEG data by using a geometric approach
called \textit{beamforming} to solve the biomagnetic or bioelectrical
inverse problem~\cite{Sarv87}.  This approach entails using detailed
brain conductance models (often informed by high spatial resolution
anatomical MRI images) along with the known sensor placements
(localized precisely in 3D space) to reconstruct brain signals
originating from theoretical point sources deep in the brain (and far
from the sensors).  Traditional beamforming approaches must overcome
two obstacles.  First, the inverse problem beamforming seeks to solve
has infinitely many solutions.  Researchers have made traction towards
constraining the solution space by assuming that signal-generating
sources are localized on a regularly spaced grid spanning the brain
and that individual sources are small relative to their distances to
the sensors~\cite{Snyd91, BailEtal01, HillEtal05}.  The second, and in
some ways much more serious, obstacle is that the magnetic fields
produced by external (noise) sources are substantially stronger than
those produced by the neuronal changes being sought (i.e., at deep
structures, as measured by sensors at the scalp).  This means that
obtaining adequate signal quality often requires averaging the
measured responses over tens to hundreds of responses or trials
(e.g., see review by~\cite{HillEtal05}).

Another approach to obtaining high spatial and temporal resolution
neural data has been to collect fMRI and EEG data simultaneously.
Simultaneous fMRI-EEG has the potential to balance the high spatial
resolution of fMRI with the high temporal resolution of scalp EEG,
thereby, in theory, providing the best of both worlds.  In practice,
however, the signal quality of both recordings suffers substantially
when the two techniques are applied simultaneously (e.g., see review
by~\cite{HustEtal12}).  In addition, the experimental designs that are
ideally suited to each technique individually are somewhat at odds.
For example, fMRI experiments typically lock stimulus presentation
events to the regularly spaced image acquisition time (TR), which
maximizes the number of post-stimulus samples.  By contrast, EEG
experiments typically employ jittered stimulus presentation times to
maximize the experimentalist's ability to distinguish electrical brain
activity from external noise sources such as from 60 Hz alternating
current power sources.

The current ``gold standard'' for precisely localizing signals and
sampling at high temporal resolution is to take (ECoG or iEEG)
recordings from implanted electrodes (but from a limited set of
locations in any given brain).  This begs the following question: what
can we infer about the activity exhibited by the rest of a person's
brain, given what we learn from the limited intracranial recordings we
have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we
call \textit{SuperEEG}, based on Gaussian process
regression~\cite{Rasm06}.  SuperEEG entails using data from multiple
people to estimate activity patterns at arbitrary locations in each
person's brain (i.e., independent of their electrode placements).  We
test SuperEEG approach using two large datasets of intracranial
recordings~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12, EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  We show
that the SuperEEG algorithm recovers signals well from electrodes that
were held out of the training dataset.  We also examine the factors
that influence how accurately activity may be estimated (recovered),
which may have important implications for electrode design and
placement in neurosurgical applications.

@@

We first applied fourth order Butterworth notch filter to remove 60 Hz
($\pm$ .5 Hz) line noise from every recording (from every electrode).
We then excluded any electrodes that showed putative epileptiform
activity.  Specifically, we excluded from further analysis any
electrode that exhibited an average kurtosis of 10 or greater across
all of that patient's recording sessions.  We also excluded any
patients with fewer than 2 electrodes that passed this criteria, as
the SuperEEG algorithm requires measuring correlations between 2 or
more electrodes from each patient.  For Dataset 1, this yielded clean
recordings from 4168 electrodes implanted throughout the brains of 67
patients (Fig.~\ref{fig:methods}A); for Dataset 2, this yielded clean
recordings from 2975 electrodes from 24 patients.  Each individual
patient contributes electrodes from a limited set of brain locations,
which we localized in a common space~\citep[MNI152;][]{GrabEtal06}; an
example Dataset 1 patient's 54 electrodes that passed the predefined
kurtosis test are highlighted in black and red.

The recording from a given electrode is maximally informative about
the activity of the neural tissue immediately surrounding its
recording surface.  However, brain regions that are distant from the
recording surface of the electrode also contribute to the recording,
albeit (ceteris paribus) to a much lesser extent.  One mechanism
underlying these contributions is volume conduction.  The precise rate
of falloff due to volume conduction (i.e., how much a small volume of
brain tissue at location $x$ contributes to the recording from an
electrode at location $\eta$) depends on the size of the recording
surface, the electrode's impedance, and the conductance profile of the
volume of brain between $x$ and $\eta$.  As an approximation of this
intuition, we place a Gaussian radial basis function (RBF) at the
location $\eta$ of each electrode's recording surface
(Fig.~\ref{fig:methods}B).  We use the values of the RBF at any brain
location $x$ as a rough estimate of how much structures around $x$
contributed to the recording from location $\eta$:
\begin{align}
  \mathrm{rbf}(x|\eta,\lambda) & =
  \mathrm{exp}\left\{ -\frac{||x - \eta||^2}{\lambda} \right\},\label{eqn:rbf}
\end{align}
where the width variable $\lambda$ is a parameter of the algorithm
(which may in principle be set according to location-specific tissue
conductance profiles) that governs the level of spatial smoothing.  In
choosing $\lambda$ for the analyses presented here, we sought to
maximize spatial resolution (which implies a small value of $\lambda$)
while also maximizing the algorithm's ability to generalize to any
location throughout the brain, including those without dense electrode
coverage (which implies a large value of $\lambda$). Using our prior
work as a guide~\citep{MannEtal14b, MannEtal18}, we set $\lambda = 20$,
although this could in theory be optimized, e.g., using cross
validation or a formal model~\citep[e.g.,][]{MannEtal18}.

A second mechanism whereby a given region $x$ can contribute to the
recording at $\eta$ is through anatomical connections between
structures near $x$ and $\eta$.  We use spatial correlations in the
data to estimate these anatomical connections~\citep{BeckEtal18}.  Let
$\bar{R}$ be the set of locations at which we wish to estimate local
field potentials, and let $R_{s}$ be set of locations at which we
observe local field potentials from patient $s$ (excluding the
electrodes that did not pass the kurtosis test described above). In
the analyses below we define $\bar{R} = \cup_{s=1}^S R_s$.  We can
calculate the expected inter-electrode correlation matrix for patient
$s$, where $C_{s,k}(i,j)$ is the correlation between the time series
of voltages for electrodes $i$ and $j$ from subject $s$ during session
$k$, using:

\begin{align}
  \bar{C}_{s} &=
  \mathrm{r}(\frac{1}{n}(\sum_{k=1}^{n}\mathrm{z}(C_{s,k}))),\label{eqn:inter_corr}~\mathrm{where}\\
\mathrm{z}(r) &= \frac{\log(1+r) - \log(1 - r)}{2}~\mathrm{is~the~Fisher}~z \mathrm{-transformation~and}\label{eqn:fishersz}\\
\mathrm{z}^{-1}(z) = \mathrm{r}(z) &= \frac{\exp(2z) - 1}{\exp(2z) + 1}\mathrm{~is~its~inverse}.\label{eqn:invfishersz}
\end{align}
Next, we use Equation~\ref{eqn:rbf} to construct a number of
to-be-estimated locations by number of patient electrode locations
weight matrix, $W_s$.  Specifically, $W_s$ approximates how informative the
recordings at each location in $R_s$ are in reconstructing activity at
each location in $\bar{R}$, where the contributions fall off with an RBF according to
the distances between the corresponding locations:

\begin{align}
W_s(i, j) = \mathrm{rbf}(i|j,\lambda)\label{eqn:weight_matrix}.
\end{align}

Given this weight matrix, $W_s$,
and the observed inter-electrode correlation matrix for patient $s$,
$\bar{C}_{s}$, we can estimate the correlation matrix for all locations in
$\bar{R}$ ($\hat{C}_s$; Fig.~\ref{fig:methods}C) using:

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:subj_corrmat_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i)
                    \cdot W(y,j). \label{eqn:subj_corrmat_den}\\
 \hat{C}_s &= \mathrm{r}\left( \frac{\hat{N}_{s}}{\hat{D}_{s}} \right)
             \label{eqn:subj_corrmat}.
\end{align}
After estimating the numerator ($\hat{N}_{s}$) and denominator
($\hat{D}_{s}$) placeholders for each $\hat{C}_{s}$, we aggregate these
estimates across patients to obtain a single expected full-brain
correlation matrix ($\hat{K}$; Fig.~\ref{fig:methods}D):
\begin{align}
 \hat{K} = \mathrm{r} \left(  \frac{\sum  \hat{N}_{s}}{\sum \hat{D}_{s}}\right).\label{eqn:correlation_matrix}
\end{align}
Intuitively, the numerators capture the general structures of the
patient-specific estimates of full-brain correlations, and the
denominators account for which locations were near the implanted
electrodes in each patient.  To obtain $\hat{K}$, we compute a
weighted average across the estimated patient-specific full-brain
correlation matrices, where patients with observed electrodes near a
particular set of locations in $\hat{K}$ contribute more to the
estimate.

Having used the multi-patient data to estimate a full-brain
correlation matrix at the set of locations in $\bar{R}$ that we wish
to know about, we next use $\hat{K}$ to estimate activity patterns
everywhere in $\bar{R}$, given observations at only a subset of
locations in $\bar{R}$ (Fig.~\ref{fig:methods}E).

Let $\alpha_s$ be the set of indices of patient $s$'s electrode locations in
$\bar{R}$, and let $\beta_s$ be the set of indices of all other
locations in $\bar{R}$. In other words, $\beta_s$ reflects the locations
in $\bar{R}$ where we did not observe a recording for patient $s$
(these are the recording locations we will want to fill in using SuperEEG).
We can sub-divide $\hat{K}$ as follows:

\begin{align}
\hat{K}_{\beta_s,\alpha_s} &= \hat{K}(\beta_s,\alpha_s),~\mathrm{and}\label{eqn:Kba}\\
\hat{K}_{\alpha_s,\alpha_s} &= \hat{K}(\alpha_s,\alpha_s)\label{eqn:Kaa}.
\end{align}
Here $\hat{K}_{\beta_s, \alpha_s}$ represents the correlations between
the ``unknown'' activity at the locations in $\beta_s$ and the
observed activity at the locations in $\alpha_s$, and
$\hat{K}_{\alpha_s, \alpha_s}$ represents the correlations between the
observed recordings (at the locations in $\alpha_s$).

Let $Y_{s,k,\alpha_s}$ be the number-of-timepoints ($T$) by
length($\alpha_s$) matrix of (observed) voltages from the electrodes in
$\alpha_s$ during session $k$ from patient $s$. Then we can estimate the
voltage from patient $s$'s $k^{th}$ session at the locations in
$\beta_s$ using~\cite{Rasm06}:

\begin{align}
Y_{s,k,\beta} &= ((\hat{K}_{\beta_s,\alpha_s}\cdot\hat{K}_{\alpha_s,\alpha_s}^{-1})\cdot Y_{s,k,\alpha_s}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas we
observe recordings only at the locations in $\alpha_s$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations in $\beta_s$, which we can define \textit{a priori}
to include any locations we wish, throughout the brain.  This yields
estimates of the time-varying voltages at \textit{every} location in
$\bar{R}$, provided that we define $\bar{R}$ in advance to include the
union of all of the locations in $\alpha_s$ and all of the locations
at which we wish to estimate recordings (i.e., a timeseries of
voltages).

We designed our approach to be agnostic to electrode impedances, as
electrodes that do not exist do not have impedances.  Therefore our
algorithm recovers voltages in standard deviation ($z$-scored) units
rather than attempting to recover absolute voltages. (This property
reflects the fact that $\hat{K}_{\beta_s, \alpha_s}$ and
$\hat{K}_{\alpha_s, \alpha_s}$ are correlation matrices rather than
covariance matrices.)  Also, note that
Equation~\ref{eqn:reconstruction} directly requires computing a $T$ by $T$
matrix, which can become computationally intractable when $T$ is very
large (e.g., for the patient highlighted in Fig.~\ref{fig:recon}, $T =
20458799$). However, because Equation~\ref{eqn:reconstruction} is time
invariant, we may compute $Y_{s,k,\beta_s}$ in a piecewise
manner by filling in $Y_{s,k,\beta_s}$ one row at a time (using the
corresponding samples from $Y_{s, k, \alpha_s}$).

The SuperEEG algorithm described above and in
Figure~\ref{fig:methods} allows us to estimate (up to a constant
scaling factor) LFPs for each patient at all arbitrarily chosen
locations in the set $\bar{R}$, \textit{even if we did not record that
  patient's brain at all of those locations}.  We next turn to an
evaluation of the accuracy of those estimates.

@@

@@

Holding out each electrode from each patient in turn, we computed the
average correlation (across recording sessions) between the Super
EEG-reconstructed voltage traces and the observed voltage traces from
that electrode.  For each reconstruction, we estimated the full-brain
correlation matrix using every \textit{other} patient's data
(i.e., every patient except the one who contributed the
to-be-reconstructed electrode data).  In our analyses, we then
substituted the average correlation matrix computed after excluding
patient $s$'s data for $\hat{K}$ in Equations~\ref{eqn:Kba} and
\ref{eqn:Kaa}.  This step ensured that the data we were reconstructing
could not also be used to estimate the between-location correlations
that drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).

@@

%why does it work?  redundancy of brain
A fundamental assumption of the SuperEEG algorithm is that the data
covariance matrix is stable over time and across people.  This is a
useful simplification.  However, a growing body of evidence from the
fMRI community suggests that the data covariance matrix changes in
meaningful ways over time (for example, the data covariance matrix
changes from moment-to-moment during story listening, serving as a
unique ``fingerprint'' for each moment of the story; further, these
task-driven timepoint-specific covariance fingerprints appear to be
largely preserved across people~\cite{SimoEtal16, MannEtal17}).  These
findings indicate that the full-brain covariance matrix is not stable
over time.  Other recent work has shown that people's resting state
connectivity matrices may be used to uniquely identify individuals and
predict fluid intelligence scores~\cite{FinnEtal15}.  This indicates
that the full-brain covariance matrix is not stable across people.  If
the fundamental stability assumptions that SuperEEG relies on are
violated, how can the SuperEEG algorithm still accurately recover LFP
data?  It is important to recognize that the fact that variability
(over time or across people) is predictive (e.g., of cognitive states
during story listening or fluid intelligence scores) does not
necessarily mean that this variability is large in magnitude.  Rather,
we have long known that brain structure is tightly preserved across
individuals (and over time, at least on the timescale of typical
clinical and experimental recording sessions), and any functional
changes must occur within the framework of the underlying structural
anatomy.  Nevertheless, one could imagine future improvements to the
SuperEEG approach that leverage resting state fMRI or structural data
[e.g., diffusion tensor imaging (DTI)] to estimate Bayesian priors over
the correlation matrices inferred, in the current framing, using only
ECoG data.  Further, relaxing the assumption that the covariance
matrix is stable (over time and/or across people), and/or
incorporating more detailed brain conductance models (e.g., informed
by structural MRI scans) may improve the predictive performance of the
approach.
